---
title: "day-10-afternoon"
author: "Daniel Kerstan"
date: "8/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

```{r}
meteor <- read_csv(here("data", "ntl20_v6.csv"))

ice_dur <- read_csv(here("data", "ntl33_v7.csv"), na = "-999")
  
```

```{r}
ice <- ice_dur %>% 
  group_by(lakeid) %>% 
  summarize(mean_ice_duration = mean(ice_duration, na.rm = TRUE))
```

```{r}
ice_plot <- ggplot(ice_dur, aes(x = lakeid, y = ice_duration)) +
  geom_jitter() +
  geom_violin(fill = NA)



ice_plot
```

```{r}
ice_mean_year <- ice_dur %>% 
  group_by(year4) %>% 
  summarize(mean_ice_duration = mean(ice_duration, na.rm = TRUE))
ice_mean_year_plot <- ggplot(ice_mean_year, aes(x = year4, y = mean_ice_duration)) +
  geom_point()

ice_mean_year_plot

```
There is a slight negative trend showing ice duration decreasing over time

```{r}
meteor_mean_air <- meteor %>% 
  filter(month %in% c(12, 1, 2)) %>% 
  group_by(year4) %>% 
  summarize(mean_air_temp = mean(ave_air_temp_adjusted))

meteor_mean_air_plot <- ggplot(meteor_mean_air, aes(x = year4, y = mean_air_temp)) +
  geom_point() +
  geom_line() 


meteor_mean_air_plot
```
there is a slight positive trend, where average air temperature is increasing over time


```{r}
combined <- full_join(ice_mean_year, meteor_mean_air)

combined <- combined %>% 
  drop_na(mean_air_temp)
```

```{r}
combined_plot <- ggplot(combined, aes(x = mean_air_temp, y = mean_ice_duration)) +
  geom_point()

combined_plot
```
there is a slight negative trend showing that ice duration decreases with increasing air temperature

```{r}
ice_cover <- ice_dur %>% 
  filter(year4 >= 1970) %>% 
  mutate(ice_off = ymd(ice_off)) %>% 
  mutate(ice_off_day = yday(ice_off)) %>% 
  select(year4, ice_off_day)

ggplot(ice_cover, aes(x= year4, y = ice_off_day)) +
  geom_point()
  
  
```
It is difficult to distinguish a trend from this plot. there is a possible negative trend indicating the ice off day has become earlier over time

